---
title: "Individual1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
install.packages('readxl')
data <- readxl::read_excel('/Users/visma/Desktop/DS202/ds202_lab3/GSS.xls')
data
```

Question 1

```{r}
years <- as.numeric(data$`Highest year of school completed`)
data$years <- years
var1 <- data %>%
  group_by(`Marital status`) %>%
    summarize(AvgYr = mean(years, na.rm = TRUE)) %>%
      arrange(AvgYr)

var1
ggplot(data = var1, aes(x = AvgYr, y = `Marital status`)) + geom_bar(stat="identity")

```

There are no explicit patterns discernible from the data, but it can be claimed that people who never married had on average more years of education that people who did marry , regardless of how the relationship ended.


Question 2

```{r}
##For No Answer
nAns <- data.frame(data[data$`Marital status` %in% c('No answer'), ])
nAns<- na.omit(nAns)
AvgNAns <- mean(nAns$years)
FNAns <- nAns[!nAns$years < AvgNAns, ]

##For Never Married
nm <- data.frame(data[data$`Marital status` %in% c('Never married'), ])
nm<- na.omit(nm)
AvgNM <- mean(nm$years)
FNM <- nm[!nm$years < AvgNM, ]

##For Divorced
d <- data.frame(data[data$`Marital status` %in% c('Divorced'), ])
d<- na.omit(d)
AvgD <- mean(d$years)
FD <- d[!d$years < AvgD, ]

##For Married
m <- data.frame(data[data$`Marital status` %in% c('Married'), ])
m<- na.omit(m)
AvgM <- mean(m$years)
FM <- m[!m$years < AvgM, ]

##For Separated
s <- data.frame(data[data$`Marital status` %in% c('Separated'), ])
s<- na.omit(s)
AvgS <- mean(s$years)
FS <- s[!s$years < AvgS, ]

##For Widowed
w <- data.frame(data[data$`Marital status` %in% c('Widowed'), ])
w<- na.omit(w)
AvgW <- mean(w$years)
FW <- w[!w$years < AvgW, ]

FinalDF <- rbind(FNAns, FNM, FD, FM, FS, FW)
FinalDF
```


Question 3
```{r}
happy <- filter(data, `General happiness` == "Pretty happy" | `General happiness` == "Very Happy")
ggplot(happy, aes(x = `Marital status`)) + geom_histogram(stat = "count")

data %>% 
  select(`General happiness`, `Marital status`) %>%
    group_by(`Marital status`) %>%
      count(`General happiness`)
```

We can see from the graphs above that people who are married tend to be happier than people who are widowed or those with failed relationships. The second happiest people are those who never married, followed by those that are divorced. The least happiest people are the ones who are separated.


Question 4

```{r}
ggplot(data, aes(x = `Marital status`)) + geom_bar() + facet_wrap(~`Political party affiliation`) + coord_flip()

df <- data %>%
  select(`General happiness`, `Political party affiliation`, `Marital status`) %>%
    group_by(`Political party affiliation`, `Marital status`) %>%
      count(`General happiness`)

df
```

From the above we can see that political affiliation doesn't affect the relationship found in question three. This is because the distribution of happy couples is relatively proportional across all party affiliations. In every party, married couples are the happiest by far, and separated are the least.